---
title: "Untitled"
author: "Surya Vamshidhar Buneeti"
date: "2024-04-08"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ISLR) 
library(caret) 
```
```{r}
library(dplyr) 
```
```{r}
library(cluster) 
library(factoextra)
```
```{r}
library(NbClust) 
library(ppclust) 
library(dendextend) 
```
```{r}
library(tidyverse)
```
```{r}
library(ggplot2)
library(proxy)

data <- read.csv("C:/Users/surya/Downloads/Cereals.csv")
#printing some records form the dataset head(CEREALS_DATA) 


```

```{r}
#printing the summary of the dataset summary(CEREALS_DATA) 
```


```{r}
#initializing a dataframe to store the pre-processsed data cereal_scaled_data <- CEREALS_DATA 
#scaling the cereals data and preparing for the clustering algorithm cereal_scaled_data[ , c(4:16)] <- scale(CEREALS_DATA[ , c(4:16)]) 
#removing unwanted null values cleansed_data <- na.omit(cereal_scaled_data) 
#printing the pre-processed data head(cleansed_data) 

```



```{r}
#calculating the euclidean distance 
Euclidean_dist <- dist(Preprocessed_Cereals_data[ , c(4:16)], method = "euclidean") 
#Euclidean_dist	 

```

```{r}
#using AGNES to compare and single linkage method for clustering 
clustering_H_single <- agnes(Euclidean_dist, method = "single")	 

```
 
 
```{r}
plot(clustering_H_single, 
main = "Ratings, AGNES comparision (Single Linkage)",      xlab = "Height",      ylab = "Cereal",      cex.axis = 0.5,      cex = 0.5,      hang = -1) 
```
```{r}
#applying AGNES with complete linkage for clustering 
clustering_H_complete <- agnes(Euclidean_dist, method = "complete")	 

```


plotting the clustering with complete linkage method 
```{r}
plot(clustering_H_complete,  
     main = "Ratings AGNES comparision (Complete Linkage)",      xlab = "Height",      ylab = "Cereal",      cex.axis = 0.5,      cex = 0.50,      hang = -1) 

```
```{r}
#applying AGNES with Average linkage method 
clustering_H_average <- agnes(Euclidean_dist, method = "average")	 

```

```{r}
plot(clustering_H_average,  
     main = "Ratings,  AGNES comparision (Average Linkage)",      xlab = "Height",      ylab = "cereal",      cex.axis = 0.5,      cex = 0.50,      hang = -1) 

```


```{r}
clustering_H_WARD <- agnes(Euclidean_dist, method = "ward") 
```



```{r}
plot(clustering_H_WARD,  
     main = "Ratings, AGNES comparision (Ward Linkage)",      xlab = "Height",      ylab = "cereal",      cex.axis = 0.5,      cex = 0.50,      hang = -1) 

```
If the value is near to 1.0, the clustering structure is closer. The approach that comes the closest to 1.0 will be selected as a result. 
Linkage Solely: 0.61 
Total Connection: 0.84 
Linkage on average: 0.78 
Ward Approach: 0.90 
Here The Ward approach is the most effective clustering model based on the results.

```{r}
 
# using the fviz_nbclust() function from the factoextra package 
fviz_nbclust(Preprocessed_Cereals_data[ , c(4:16)], hcut, method = "silhouette", k.max = 
30) + 

  labs(title = "Calculating Optimal Number of Clusters using silhouette Method") 

```
Based on the elbow and silhouette technique results, we can determine that 12 clusters would be the ideal quantity.

```{r}
fviz_nbclust(Preprocessed_Cereals_data[ , c(4:16)], hcut, method = "wss", k.max = 30) +   labs(title = "Calculating Optimal Number of Clusters using Elbow Method") +    geom_vline(xintercept = 12, linetype = 2) 
```

```{r}
plot(clustering_H_WARD,  
     main = "Clustering(12) AGNES using Ward linkage",      xlab = "Height",      ylab = "Cereal",      cex.axis = 0.5,      cex = 0.50,      hang = -1)
rect.hclust(clustering_H_WARD, k = 12, border = 1:12)

```

QUE:-The elementary public schools would like to choose a set of Cereals to include in their daily cafeterias. Every day a different cereal is offered, but all Cereals should support a healthy diet. For this goal, you are requested to find a 
cluster of “healthy Cereals.” Should the data be normalized? If not, how should they be used in the cluster analysis?


ANS:-Since the nutritional information for cereal is based on the sample of cereal under evaluation, normalizing the data would not be suitable in this situation. As a result, the data collection could only include cereals with a very high sugar content and very little iron, fiber, or other nutritional information. Once cereal has been normalized throughout the sample set, it is hard to forecast how much nutrients a child will get from it. It is plausible to infer that a cereal with an iron content of 0.999 provides almost all of the necessary iron for a child's diet; yet, it is also plausible that this cereal represents the exception rather than the rule and is nutritionally worthless.Consequently, Preprocessing the data into a ratio of a child's daily recommended intake of calories, fiber, carbohydrates, and other nutrients would be a preferable approach. This would allow analysts to make better-informed cluster judgments during the review process and prevent a few larger variables from overriding the distance estimations. When analyzing the clusters, an analyst may use the cluster average to determine the percentage of a student's daily nutritional needs that XX cereal would meet. Workers would be equipped with expertise to select "healthy" cereal clusters as a result.
